Indecision under time pressure arises from suboptimal switching behaviour

### 2024-11-05 paper/Cover_letter_Indecisions.tex
updated for jep human perception and performance

\begin{document}
\fontsize{12pt}{12pt}\selectfont
%\pagenumbering{gobble}
\pagestyle{fancy}
\cfoot{}
\lfoot{}
\rfoot{\textcolor{myudblue}{1}}
\raggedright
\begin{minipage}[ht]{0.6\textwidth}
\hspace{-10mm}\includegraphics[height=0.9in]{udcoe_logo}
\end{minipage}\hfill
\begin{minipage}[ht]{0.275\textwidth}
\begin{tabular}{lr}
\raggedleft
&\fontsize{10}{10}\selectfont \hspace{+20mm}\textcolor{myudblue}{Joshua Cashaback}\\
&\fontsize{10}{10}\selectfont \textcolor{myudblue}{Assistant Professor}\\
 &{\fontsize{10}{10}\selectfont \textcolor{myudblue}{\faPhone \hspace{1mm} +1.302.784.5986}}\\
 &{\fontsize{10}{10}\selectfont \textcolor{myudblue}{\href{mailto:joshcash@udel.edu}{\faEnvelope \hspace{1mm} joshcash@udel.edu}}}\\
%&{\fontsize{10}{10}\selectfont \textcolor{myudblue}{\faHome joshcashaback.weebly.com}}
%&{\fontsize{10}{10}\selectfont \textcolor{myudblue}{\href{mailto:cashabackjga@gmail.com}{cashabackjga@gmail.com}}}\\
\end{tabular}
\end{minipage}
\\
\vspace{6mm}
\today
\\
\vspace{6mm}
Dear Editor,
\justify
\vspace{1.0mm}
We are writing to request that you consider our paper, \textcolor{myudblue}{\textbf{Indecision under time pressure arises from suboptimal switching behaviour}}, for publication in \emph{Journal of Experimental Psychology: General}.
\vspace{3.0mm}
\\
Indecisions can be catastrophic, preventing a life-saving response or leading to a loss of large sums of money.  Despite their importance and high prevalence, indecisions have been surprisingly ignored in the literature. The main culprit is that the field of decision-making leans heavily on the classic two-alternative forced choice task---where a human or animal \emph{must} select one of two potential options---which simply does not permit a non-response past some deadline (i.e., indecision). Of the few papers that have considered a deadline, indecisions are usually only mentioned in passing and, more often, simply removed from analyses and discarded from model fitting. \textbf{The field has consequently overlooked a common and hallmark feature of human and animal behaviour---indecisions.}
\vspace{3mm}
\\
Here we designed a novel paradigm to study high pressure scenarios that lead to indecisions. We found a median indecision rate of approximately 20\%, \textbf{with some participants even displaying over 50\% indecisions despite having plenty of time to make a guess.} Our computational modelling suggested that these indecisions arose by not accounting for the time delay and uncertainty associated when switching from reacting to guessing. In a second experiment, we show for the first time there is an inherent and functionally relevant time delay when switching from reacting to guessing, a phenomenon previously unreported in the literature. \textbf{Our paper novelly provides a mechanism to explain how indecisions arise.}
\vspace{3mm}
\\
We fully expect that our work will broadly draw interest from fields as diverse as Decision-Making, Economics, and Sports Science, which are all interested in the role of decision-making on behaviour.
Thank you for your time and consideration.
\vspace{3mm}
\\
Sincerely, 
\vspace{3mm}
\\
\includegraphics[height=0.75in]{signature}
\\
\vspace{-3.5mm}
\\
Joshua G. A. Cashaback, Ph.D.
\thispagestyle{fancy}
\rfoot{\textcolor{myudblue}{\thepage}}
\lfoot{\textcolor{myudblue}{\textbf{\fontsize{10}{10}\selectfont Biomedical Engineering | College of Engineering | University of Delaware} }
\\
\fontsize{10}{10}\selectfont \textcolor{myudblue}{STAR Health Sciences Complex, Rm 201J | 540 South College Avenue | Newark, DE., 19713}}
\newpage
\noindent \textcolor{myudblue}{REVIEWER SUGGESTIONS}
\\
\\
Laurence Maloney, New York University, laurence.maloney@nyu.edu

>>>>

\begin{document}
\fontsize{12pt}{12pt}\selectfont
%\pagenumbering{gobble}
\pagestyle{fancy}
\cfoot{}
\lfoot{}
\rfoot{\textcolor{myudblue}{1}}
\raggedright
\begin{minipage}[ht]{0.6\textwidth}
\hspace{-10mm}\includegraphics[height=0.9in]{udcoe_logo}
\end{minipage}\hfill
\begin{minipage}[ht]{0.275\textwidth}
\begin{tabular}{lr}
\raggedleft
&\fontsize{10}{10}\selectfont \hspace{+20mm}\textcolor{myudblue}{Joshua Cashaback}\\
&\fontsize{10}{10}\selectfont \textcolor{myudblue}{Assistant Professor}\\
 &{\fontsize{10}{10}\selectfont \textcolor{myudblue}{\faPhone \hspace{1mm} +1.302.784.5986}}\\
 &{\fontsize{10}{10}\selectfont \textcolor{myudblue}{\href{mailto:joshcash@udel.edu}{\faEnvelope \hspace{1mm} joshcash@udel.edu}}}\\
%&{\fontsize{10}{10}\selectfont \textcolor{myudblue}{\faHome joshcashaback.weebly.com}}
%&{\fontsize{10}{10}\selectfont \textcolor{myudblue}{\href{mailto:cashabackjga@gmail.com}{cashabackjga@gmail.com}}}\\
\end{tabular}
\end{minipage}
\\
\vspace{6mm}
\today
\\
\vspace{6mm}
Dear Editor,
\justify
\vspace{1.0mm}
We are writing to request that you consider our paper, \textcolor{myudblue}{\textbf{Indecision under time pressure arises from suboptimal switching behaviour}}, for publication in \emph{Journal of Experimental Psychology: Human Perception and Performance}.
\vspace{3.0mm}
\\
Indecisions can be catastrophic, preventing a life-saving response or leading to a loss of large sums of money.  Despite their importance and high prevalence, indecisions have been surprisingly ignored in the literature. The main culprit is that the field of decision-making leans heavily on the classic two-alternative forced choice task---where a human or animal \emph{must} select one of two potential options---which simply does not permit a non-response past some deadline (i.e., indecision). Of the few papers that have considered a deadline, indecisions are usually only mentioned in passing and, more often, simply removed from analyses and discarded from model fitting. \textbf{The field has consequently overlooked a common and hallmark feature of human and animal behaviour---indecisions.}
\vspace{3mm}
\\
Here we designed a novel paradigm to study high pressure scenarios that lead to indecisions. We found a median indecision rate of approximately 20\%, \textbf{with some participants even displaying over 50\% indecisions despite having plenty of time to make a guess.} Our computational modelling suggested that these indecisions arose by not accounting for the time delay and uncertainty associated when switching from reacting to guessing. In a second experiment, we show for the first time there is an inherent and functionally relevant time delay when switching from reacting to guessing, a phenomenon previously unreported in the literature. \textbf{Our paper novelly provides a mechanism to explain how indecisions arise.}
\vspace{3mm}
\\
We fully expect that our work will broadly draw interest from fields as diverse as Decision-Making, Economics, and Sports Science, which are all interested in the role of decision-making on behaviour.
Thank you for your time and consideration.
\vspace{3mm}
\\
Sincerely, 
\vspace{3mm}
\\
\includegraphics[height=0.75in]{signature}
\\
\vspace{-3.5mm}
\\
Joshua G. A. Cashaback, Ph.D.
\thispagestyle{fancy}
\rfoot{\textcolor{myudblue}{\thepage}}
\lfoot{\textcolor{myudblue}{\textbf{\fontsize{10}{10}\selectfont Biomedical Engineering | College of Engineering | University of Delaware} }
\\
\fontsize{10}{10}\selectfont \textcolor{myudblue}{STAR Health Sciences Complex, Rm 201J | 540 South College Avenue | Newark, DE., 19713}}
\newpage
\noindent \textcolor{myudblue}{REVIEWER SUGGESTIONS}
\\
\\
Laurence Maloney, New York University, laurence.maloney@nyu.edu

### 2024-11-05 paper_apa/main.tex
removed supplementary from main pdf, and had to adjust things for the supp itself to get it to run
Update spacing
updated for jep human perception and performance

\newpage
% \pagenumbering{arabic} % page numbering starts here
Word count: 7114
\section{Introduction}
\vspace{2mm}
\noindent Indecisions arise from failing to decide and act upon sensory information in time, such as a driver failing to brake or hit the gas pedal when a traffic light turns yellow. When acting under high pressure time constraints, the ability to accurately time a decision is critical to success. The vast majority of decision-making research either does not consider responses made after some time constraint or simply does not permit a non-response, such as in the classic two-alternative forced choice paradigm \autocite{zacksenhouseRobustOptimalStrategies2010, bogaczPhysicsOptimalDecision2006, choMechanismsUnderlyingDependencies2002, joganNewTwoalternativeForced2014, ratcliffDiffusionDecisionModel2008,ulrichThresholdEstimationTwoalternative2004,tylerSignalDetectionTheory2000}. Thus, despite its real-world ubiquity and importance, we have very little understanding of how indecisions arise.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  RESULTS  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{2mm}
\section{Results}
\vspace{-1mm}

\subsection{Experiment 1}
\vspace{-1mm}

@@

\noindent 44 participants participated across two experiments. 20 individuals participated in \boldblue{Experiment 1} and 24 individuals participated in \boldblue{Experiment 2}. All participants reported they were free from musculoskeletal injuries, neurological conditions, or sensory impairments. In addition to a base compensation of \$5.00, we informed them they would receive a performance-based compensation of up to \$5.00. Participants received the full \$10.00 once they completed the experiment irrespective of their performance. All participants provided written informed consent to participate in the experiment and the procedures were approved by the University of Delaware’s Institutional Review Board.

\subsection{Apparatus}

@@

\noindent We used analysis of variance (ANOVA) as omnibus tests to determine whether there were main effects and interactions. We report the Greenhouse-Geiser adjusted p-values and degrees of freedom. In \boldblue{Experiment 1} we used a 3 (mean: early, middle, late) x 2 (variance: low, high) repeated measures ANOVA for each dependent variable. In \boldblue{Experiment 2} we used a 2 (condition: interleaved react and guess, \emph{react or guess} only) x 2 (trial type: react trials, guess trials) repeated measures ANOVA for each dependent variable. For both \boldblue{Experiment 1} and \boldblue{2}, we performed mean comparisons using nonparametric bootstrap hypothesis tests (n = 1,000,000) \autocite{hesterbergBootstrap2011,cashabackDissociatingErrorbasedReinforcementbased2017,cashabackGradientReinforcementLandscape2019a}. Mean comparisons were Holm-Bonferonni correct to account for multiple comparisons. Significance threshold was set at $\alpha$ = 0.05). We also report the common-language effect size ($\hat{\theta}$).

@@

\newpage
\input{supplementary}

>>>>

\newpage
% \pagenumbering{arabic} % page numbering starts here
\section{Introduction}
\vspace{2mm}
\noindent Indecisions arise from failing to decide and act upon sensory information in time, such as a driver failing to brake or hit the gas pedal when a traffic light turns yellow. When acting under high pressure time constraints, the ability to accurately time a decision is critical to success. The vast majority of decision-making research either does not consider responses made after some time constraint or simply does not permit a non-response, such as in the classic two-alternative forced choice paradigm \autocite{zacksenhouseRobustOptimalStrategies2010, bogaczPhysicsOptimalDecision2006, choMechanismsUnderlyingDependencies2002, joganNewTwoalternativeForced2014, ratcliffDiffusionDecisionModel2008,ulrichThresholdEstimationTwoalternative2004,tylerSignalDetectionTheory2000}. Thus, despite its real-world ubiquity and importance, we have very little understanding of how indecisions arise.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  RESULTS  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \vspace{2mm}
\section{Results}
\vspace{-2mm}

\subsection{Experiment 1}
\vspace{-2mm}

@@

\noindent 44 participants participated across two experiments (collected in 2023). 20 individuals participated in \boldblue{Experiment 1} (mean age = 25.7, 8 female) and 24 individuals participated in \boldblue{Experiment 2} (mean age = 25.1, 9 female). All participants reported they were free from musculoskeletal injuries, neurological conditions, or sensory impairments. In addition to a base compensation of \$5.00, we informed them they would receive a performance-based compensation of up to \$5.00. Participants received the full \$10.00 once they completed the experiment irrespective of their performance. All participants provided written informed consent to participate in the experiment and the procedures were approved by the University of Delaware’s Institutional Review Board.

\subsection{Power Analysis}
For \boldblue{Experiment 1}, we performed a power analysis on the interaction term of a 3 x 2 repeated measures ANOVA ($\alpha$ = 0.05, $\beta$ = 0.2 [80\% power], f = 0.34). This yielded a sample size of 14, but we collected 20 participants. For \boldblue{Experiment 2}, we performed a power analysis on the interaction term of a 2 x 2 repeated measures ANOVA ($\alpha$ = 0.05, $\beta$ = 0.2 [80\% power], f = 0.41). This yielded a sample size of 15, but we collected 24 participants.
\subsection{Apparatus}

@@

\noindent We used analysis of variance (ANOVA) as omnibus tests to determine whether there were main effects and interactions. We report the Greenhouse-Geiser adjusted p-values and degrees of freedom. In \boldblue{Experiment 1} we used a 3 (mean: early, middle, late) x 2 (variance: low, high) repeated measures ANOVA for each dependent variable. In \boldblue{Experiment 2} we used a 2 (condition: interleaved react and guess, \emph{react or guess} only) x 2 (trial type: react trials, guess trials) repeated measures ANOVA for each dependent variable. For both \boldblue{Experiment 1} and \boldblue{2}, we performed mean comparisons using nonparametric bootstrap hypothesis tests (n = 1,000,000) \autocite{hesterbergBootstrap2011,cashabackDissociatingErrorbasedReinforcementbased2017,cashabackGradientReinforcementLandscape2019a}. Mean comparisons were Holm-Bonferonni correct to account for multiple comparisons. Significance threshold was set at $\alpha$ = 0.05). We also report the common-language effect size ($\hat{\theta}$).
\subsection{Transparency and Openness}
Data, analysis scripts, and computational modeling scripts are available on github (https://github.com/CashabackLab). This study and the subsequent analyses were not preregistered.

@@

% \newpage
% \input{supplementary}
